\chapter{Background}
\label{ch:background}

In this chapter the background on both the planning and optimization aspect of this thesis is presented.
The concepts and techniques discussed in this chapter are employed in the framework we propose in \autoref{ch:methodology} with the aim of optimizing for a probabilistic model that can be used to obtain plans which maximize the performance of the system that is in need of automated control.

First of all, in \autoref{sec:decision-theoretic-planning}, the chapter discusses how \acrfull{acr:dtp} deals with the central problem of planning under uncertainty using probabilistic models of a system under consideration.
In this thesis we put our attention to modeling \acrshort{acr:dtp} problems as \acrfullpl{acr:mdp} and accordingly discuss the existing algorithmic techniques for obtaining plans for automated \acrfull{acr:sdm}.
We employ these ideas in the framework we propose in \autoref{ch:methodology} to compute optimal plans from the \acrshort{acr:mdp} models learned from data, to assess the value of different models.

Then, in \autoref{sec:bayesian-optimization}, the chapter elaborates upon a method known as \acrfull{acr:bo}, which is used to automate the process of optimizing the parameters of an unknown objective while minimizing the number of function evaluations.
This method is particularly suited for the optimization task dealt with in the framework proposed in \autoref{ch:methodology}, where the parameters of model learning algorithms are adjusted so to yield maximum performance in a system's execution.
That is, in our framework we assess the value of different \acrshort{acr:mdp} models by performing expensive simulations, which is why we would like to minimize the number of these simulations.
Therefore, in this section we describe how the \acrshort{acr:bo} framework achieves this and present some of its applications to related \acrshort{acr:sdm} problems.


\input{chapters/2.background-dtp}

\newpage

\input{chapters/2.background-bo}
